---
title: "Homework3_SDS 315"
author:
  - "Nancy Nakyung Kwak (UT EID: nk24424)"
  - "GitHub link: https://github.com/nancy1404/sds315_hw3.git"
date: "`r Sys.Date()`"
header-includes:
  - \usepackage{hyperref}
output:
  pdf_document:
    toc: true
    toc_depth: 2
---

[Link to My GitHub Repository](https://github.com/nancy1404/sds315_hw3.git) \newpage

# PROBLEM 1: Five Theories on Gas Prices

## a. Direct Competition

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(mosaic)
library(readr)

# load gas prices dataset
gasprices <- read.csv("gasprices.csv")

# visualize relationship btw gas prices and direct competition 
# Theories
# a) Gas stations charge more if they lack direct competition
ggplot(gasprices, aes( x= Competitors, y = Price)) +
  geom_boxplot() +
  labs(
    title = "Gas Prices vs. Direct Competition Presence",
    caption = "* Y shows the gas stations with direct competition, while N shows those without direct competition.
    This boxplot shows the relationship between gas price for the presence of direct competition of gas stations.",
    x = "Competitors Present (Yes or No)",
    y = "Gas Price ($)"
  )

# mean diff btw gas prices for stations with and without competition
mean_price_competition <- mean(Price ~ Competitors, data = gasprices)

# display mean prices for each category (Y and N)
mean_price_competition

# bootstrap 
boot_price_competition <- do(10000) * diffmean(Price ~ Competitors, data = resample(gasprices))

# confidence interval for diff in means
confint(boot_price_competition, level = 0.95)

```

### Theory (a) Discussion

-   Claim:
    -   Gas stations charge more if they lack direct competition in sight.
-   Evidence:
    -   The mean gas price for gas stations without direct competitors was *1.88 dollars*, while for those with direct competitors the price was *1.85 dollars*. (all numbers rounded to 2 digits after decimal.) The 95% confidence interval for the difference in mean prices is [-5.5 cents, 0.78 cents]. This intervals indicate that the difference between these two groups may not be substantial since the intervals only overlap slightly with the estimated price difference in mean price with about *2.35 cents*, having some uncertainty.
-   Conclusion:
    -   While stations without direct competitors charge slightly higher prices compared to those with direct competitors, the evidence are weak to support this theory. The overlapping confidence intervals indicate this difference may not generalize to broader population. Therefore, a larger sample or additional factors might be required to confidently confirm the meaningful difference between groups.

\newpage

## b. Rich Areas (by average income per Zipcode)

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# scatterplot for gas prices vs income
ggplot(gasprices, aes(x = Income, y = Price)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  scale_x_continuous(labels = scales::label_comma()) +
  labs(
    title = "Gas Prices vs. Income Levels",
    caption = "* This scatterplot shows the relationship between gas prices and average income levels across areas.",
    x = "Income ($)",
    y = "Gas Price ($)"
  )

cor_price_income <- cor(Price ~ Income, data = gasprices)
cor_price_income

set.seed(111) # for reproducibility
# Bootstrap for price differences by income
boot_price_income <- do(10000) * cor(Price ~ Income, data = resample(gasprices))

# Confidence interval
confint(boot_price_income, level = 0.95)


```


### Theory (b) Discussion

(\* All numbers rounded to 2 digits after decimals.)

-   Claim:
    -   The gas prices are higher in the richer areas compared to those in the less richer areas.
-   Evidence:
    -  The correlation between gas prices and icome was 0.396, showing some positive relationships. The 95% confidence interval for correlation is [0.199, 0.570], and the interval lies all in positive range, suggesting areas with higher income levels tend to have higher gas prices.
-   Conclusion:
    -   Income levels and gas prices are definitely correlated. With a correlation estimate of 0.396, there is moderate but meaningful relationship between the factors even though other analytical factors would help better describe the variations in gas prices across areas. 

\newpage

## c. Stoplights

```{r, echo = FALSE, message = FALSE, warning = FALSE}

# c) Gas stations at stoplights charge more
ggplot(gasprices, aes(x= as.factor(Stoplight), y = Price )) +
  geom_boxplot()+
  labs(
    title = "Gas Prices vs. Stoplight Presence",
    caption = "* Y represents the gas stations with a stoplight in front, while N represents those without a stoplight.
    This boxplot shows the relationship between gas price for the presence of stoplight in front of gas stations.",
    x = "Stoplight Presence (Yes or No)",
    y = "Gas Price ($)"
  )

# mean price diff btw gas stations
mean_price_stoplight <- mean(Price ~ Stoplight, data = gasprices)
mean_price_stoplight

boot_price_stoplight <- do(10000) * diffmean(Price~Stoplight, data = resample(gasprices))

confint(boot_price_stoplight, level = 0.95)
```

### Theory (c) Discussion

(\* All numbers rounded to 2 digits after decimals.)

-   Claim:
    -   Gas stations located at stoplights charge higher prices compared to those without stoplights.
-   Evidence:
    -   The mean gas price for gas stations without stoplights is *1.866 dollars*, while for those with stoplights the price is *1.863 dollars*. (round to 3 digits after decimals for this case) The 95% confidence interval for the mean price without stoplights is *[1.84, 1.90] dollars*, and for those with stoplights, the price is *[1.84, 1.88] dollars*. The estimated price difference is therefore *-0.33 cents*, with a confidence interval for difference of [-0.03, 0.03] which indicates that it is hard to say there are meaningful difference between the two groups, since the intervals overlap a lot.
-   Conclusion:
    -   The evidence does not support this theory. The confidence intervals overlaps significantly and the differences for price are near zero, which shows the stoplights' presence does not affect gas prices. The difference could be because of the random sampling variation more than just a real fundamental factor.

\newpage

## d. Highway Access

```{r, echo = FALSE, message = FALSE, warning = FALSE}

# d) Gas stations with direct highway access charge more
ggplot(gasprices, aes(x = as.factor(Highway), y = Price)) +
  geom_boxplot() +
  labs(
    title = "Gas Prices vs. Direct Highway Access",
    x = "Direct Highway Access (Yes or No)",
    y = "Gas Price ($)",
    caption = "* This boxplot shows the relationship between gas price and the gas stations with/without direct highway access."
  )

# mean diff btw gas prices for highway presence
mean_price_highway <- mean(Price ~ Highway, data = gasprices)
mean_price_highway

# bootstrap 
boot_price_highway <- do(10000) * diffmean(Price ~ Highway, data = resample(gasprices))

# confidence interval 
confint(boot_price_highway, level = 0.95)

```

### Theory (d) Discussion

(\* All numbers rounded to 2 digits after decimals.)

-   Claim:
    -   Gas stations with direct highway access charges more than those without highway access.
-   Evidence:
    -   The mean gas price for stations without highway access is *1.85 dollars*, while for stations with highway access, the price is *1.90 dollars*. The 95% confidence interval for the difference in means is [0.0082, 0.0808 dollars]. The price difference is *4.57 cents*, suggesting a meaningful increase in price for gas stations with direct highway access..
-   Conclusion:
    -   The evidence seems to support this theory. The confidence intervals do not overlap, and the estimated price difference of 4.57 cents with 95% confidence indicate the gas stations seems to charge more if they have a direct highway access. Therefore, the evidence shows a meaningful difference between two groups, suggesting a real difference in this case, not just a random variation.

\newpage

## e. Shell vs. non-shell

```{r, echo = FALSE, message = FALSE, warning = FALSE}

# e) Shell charges more than non-Shell brands

# filter data set
gasprices_filtered <- gasprices |>
  mutate(ShellBrand = ifelse(Brand == "Shell", "Shell", "Non-Shell"))

# Plot boxplot comparing Shell vs. Non-Shell
ggplot(gasprices_filtered, aes ( x= ShellBrand, y = Price)) +
  geom_boxplot()+
  labs(
    title = "Gas Price by Brand (Shell or Non-Shell)",
    caption = "* This boxplot shows the gas price per each brand, especially to 
    find out if Shell specifically charges more than non-Shell brands.",
    x = "Brand (Shell or Non-Shell)",
    y = "Gas Price ($)"
  )

# Bootstrap for difference in means
boot_price_brand <- do(10000) * diffmean(Price ~ ShellBrand, data = resample(gasprices_filtered))

# confidence interval for difference in means
confint(boot_price_brand, level = 0.95)

mean_price_brand <- gasprices_filtered |>
  group_by(ShellBrand) |>
  summarize(mean_price = mean(Price, na.rm = TRUE))
mean_price_brand


```

### Theory (e) Discussion

(\* All numbers rounded to 2 digits after decimals.)

-   Claim:
    -   "Shell" charges more than all other "non-shell" gas station brands.
-   Evidence:
    -   The mean gas price for Shell gas stations is *1.884 dollars*, while for non-Shell stations --including, Chevron-Texaco, ExxonMobil, Other--, the prices are *1.856 dollars*. The estimated price difference is 2.74 cents in favor of Shell. The 95% confidence interval for Shell is *[-0.01, 0.07 dollars]* based on bootstrapping.
-   Conclusion:
    -   The evidence does support this theory, but with some room for uncertainty. Shell stations have higher mean prices, but the confidence intervals for Shell significantly overlap with those with non-Shell stations. This indicates the difference in price is not that huge to be successfully considered statistically, where the observed differences could be because of random sampling variation more than a specific effect across stations.

\newpage

# PROBLEM 2: Mercedes S-Class Vehicles

## PART A: Average Mileage of 2011 S-class 63 AMGs

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# PART A : bootstrap for avg mileage of 2011 S-class 63 AMGs

library(boot)
sclass <- read_csv("sclass.csv")

sclass_filtered <- sclass |> #filter for 2011 63 AMG
  filter(year == 2011, trim == "63 AMG")
# 116 cars found.

# set seed for reproducibility
set.seed(111)

# boostrap resampling for avg mileage
boot_mileage <- do(10000)*mean(~mileage, data = resample(sclass_filtered))

# Histogram of bootstrap distribution
ggplot(boot_mileage) +
  geom_histogram(aes(x=mean)) +
  labs(
    title = "Bootstrap Distribution of Average Mileage for 2011 S-Class 63 AMG",
    x = "Average Mileage",
    y = "Frequency"
  )

# calculate standard error (spread of sampling distribution)
sd(~mean, data = boot_mileage)

# 95% confidence level
confint(boot_mileage, level = 0.95)

```

### A) Interpretations

-   Conclusion: **With 95% confidence, the true mean mileage of 2011 S-Class 63 AMGs lies between 26251.44 and 31815.32 miles.**
-   Further Explanation: The original mean mileage is *28997.34 miles*. The bias was negligible (-2.66), showing that bootstrap estimates are closely aligned with the original mean. The standard error shows that each bootstrap resample produces a mean mileage estimate differing by about *1415.97 miles* from the original sample mean. The lower bound (26251.44 mi) is the 2.5th percentile of the bootstrap distribution, showing that only 2.5% of bootstrap samples have mean mileage lower than this value. The upper bound (31815.32 mi) represents the 97.5th percentile of the distribution, where only 2.5% of samples have a higher mean mileage.

\newpage

## PART B: 2014 S-class 550s painted black

```{r, echo = FALSE, message = FALSE, warning = FALSE}
sclass_filtered2 <- sclass |>
  filter(year == 2014, trim == "550") |>
  mutate(isBlack = color == "Black")
# sclass_filtered2 #2889 cars found

set.seed(111) #reproducibility check

# Bootstrap resampling for proportion of black cars
boot_proportion <- do(10000)*mean(~isBlack, data = resample(sclass_filtered2))

# histogram of bootstrap distribution
ggplot(boot_proportion) +
  geom_histogram(aes(x=mean)) +
  labs(
    title = "Bootstrap Distribution of Proportion of Black 2014 S-Class 550 Cars",
    x = "Proportion of Black Cars",
    y = "Frequency"
  )

# standard error (spread of sampling distribution)
sd(~mean, data = boot_proportion)

# 95% confidence level
confint(boot_proportion, level = 0.95)

```

### B) Interpretations

-   Conclusion: **With 95% confidence, the true proportion of 2014 S-Class 550s that are painted black lies between 0.4164 and 0.4528.**
-   Further Explanation: The original proportion of black cars in the sample is *0.4348*, meaning that about 43.48% of the 2014 S-class 550s in the dataset are black. The bias is *7.41e-06*, which is very close to zero, indicating there is very small or negligible bias in this estimates. The SD is *0.0092* which again each bootstrap sample's proportion differs by about 0.92% from the original sample proportion. The lower bound of confidence level is *0.4164* which means that 2.5% of the bootstrap samples have a proportion of black cars lower than this value. The upper bound of confidence interval is *0.4528*, meaning that 2.5% of bootstrap samples have a proportion of black cars higher than this value.

\newpage

# PROBLEM 3: TV Network Survey

## PART A: Who Makes People Happier, Ed or Earl?

```{r, echo = FALSE, message = FALSE, warning = FALSE}

tvsurvey <- read.csv('nbc_pilotsurvey.csv')

# filter for shows "Living with Ed" and "My Name is Earl"
filtered_tvsurvey <- tvsurvey |>
  filter(Show %in% c("Living with Ed", "My Name is Earl"))

set.seed(111) # seed for reproducibility

# bootstrap resampling for difference in happiness mean btw two shows
boot_diff_mean = do(10000) * diffmean(Q1_Happy ~ Show, data = resample(filtered_tvsurvey))

# bootstrap sampling distribution
ggplot(boot_diff_mean) +
  geom_histogram(aes(x=diffmean)) +
  labs(
    title = "Bootstrap Distribution of Difference in Mean Happiness Ratings",
    caption = "* including Living with Ed and My Name is Earl",
    x = "Difference in Mean Happiness Ratings",
    y = "Frequency"
  )

# calculate standard error
sd(~diffmean, data = boot_diff_mean)

# 95% confidence interval for diff in means
confint(boot_diff_mean, level = 0.95)


```

### A) Conclusion

-   Question: Which show "Living with Ed" or "My Name is Earl", made people happier, and is there evidence that one show consistently produces higher mean of Q1_Happy response?
-   Approach: First I filtered data just for "Living with Ed" and "My Name is Earl" TV show, and used bootstrapping to estimate the difference in mean Q1_Happy responses between these two shows. Then I constructed a 95% confidence interval.
-   Results: The bootstrapped 95% confidence interval for the difference in mean Q1_Happy responses between the two shows is (-0.3927, 0.1028), with the original point estimate of -0.1491.
-   Conclusion: The confidence interval includes zero, which means there is no strong evidence to conclude that one show consistently produces higher happiness ratings. The original point estimate is -0.1491, suggesting that, on average, Living with Ed viewers may have rated their happiness slightly lower. However, since the confidence interval includes zero, the difference is likely due to random variation rather than a meaningful distinction between the two shows.

\newpage

## PART B: The Bigger Loser vs. The Apprentice: Los Angeles

```{r, echo = FALSE, message = FALSE, warning = FALSE}

# filter for two shows only
filtered_tvsurvey2 <- tvsurvey |>
  filter(Show %in% c("The Biggest Loser", "The Apprentice: Los Angeles"))

set.seed(111) # reproducibility
# bootstrap resampling for difference in annoyance mean btw two shows
boot_diff_mean_annoyed = do(10000) * diffmean(Q1_Annoyed ~ Show, data = resample(filtered_tvsurvey2)) 

# Bootstrap sampling distribution
ggplot(boot_diff_mean_annoyed) +
  geom_histogram(aes(x=diffmean)) +
  labs(
    title = "Bootstrap Distribution of Difference in Mean Annoyance Ratings",
    caption = "* including The Biggest Loser and The Apprentice: Los Angeles",
    x = "Difference in Mean Annoyance Ratings",
    y = "Frequency"
  )

# calculate sd error
sd(~diffmean, data = boot_diff_mean_annoyed)

# 95% confidence interval for diff in means
confint(boot_diff_mean_annoyed, level = 0.95)

```

### B) Conclusion

-   Question: Which reality/contest show, "The Biggest Loser" or "The Apprentice: Los Angeles," made people feel more annoyed, and is there evidence that oen show consistetnly produces higher mean Q1_Annoyed response?
-   Approach: First I filtered the data for just the show "The Biggest Loser" and "The Apprentice: Los Angeles", and used bootstrapping to estimate difference in mean Q1_Annoyed responses between these two shows. I then constructed 95% confidence interval for the mean difference.
-   Results: The bootstrapped 95% confidence interval for the difference in mean Q1_Annoyed responsed between "The Biggest Loser" and "The Apprentice:Los Angeles" is (-0.5193, -0.0116), with the original estimate of -0.27. The standard error of the bootstrapped mean difference is 0.1280772.
-   Conclusion: The confidence interval is entirely below zero, showing that "The Apprentice: Log Angeles" consistently produces higher annoyance ratings than "The Biggest Loser". The point estimate of -0.271 further supports this conclusion, showing that "The Biggest Loser" is rated as 0.271 points less annoying on average compared to "The Apprentice: Los Angeles". Therefore there is strong evidence to suggest that "The Apprentice: Los Angeles" made viewers feel more annoyed.

\newpage

## PART C: Dancing with the Stars, Confusing or Not?

```{r, echo = FALSE, message = FALSE, warning = FALSE}

# filter data for Dancing with the Stars show only
dancing_data <- tvsurvey |>
  filter(Show == "Dancing with the Stars") |>
  select(Q2_Confusing)

# create binary variable for confusing response 

dancing_data$Confusing <- ifelse(dancing_data$Q2_Confusing >= 4, 1, 0)

set.seed(111) #reproducibility 

# Bootstrap resampling for the proportion of confusing responses
boot_confusing <- do(10000) * mean(~Confusing, data = resample(dancing_data))

# ggplot
ggplot(boot_confusing) +
  geom_histogram(aes(x=mean), binwidth = 0.01) +
  labs(
    title = "Bootstrap distribution of Proportion of Confusing Responses",
    x = "Proportion of Confusing Responses",
    y = "Frequency"
  )


# calculate standard error (spread of sampling distribution)
sd(~mean, data = boot_confusing)

# 95% confidence interval
# quantile(boot_confusing$X1, probs= c(0.025, 0.975))
confint(boot_confusing, level = 0.95)
```

### C) Conclusion

-   Question: What proportion of American TV watchers would give a response of 4 or 5 to the "Q2_Confusing" question for the show "Dancing with the Stars"?
-   Approach: I used bootstrapping to estimate the proportion of viewers who think the show confusing, then constructed a 95% confidence interval for this proportion.
-   Results:The 95% Confidence Interval for the proportion of viewers who found "Dancing with the Starts" confusing is (0.0387, 0.1160), with an original point estimate of 0.077. The standard error was calculated as 0.0199.
-   Conclusion: Between 3.87% and 11.60% of viewers are likely to think the show "Dancing with the Stars" confusing. The original point estimate 0.077 shows around 7.7% viewers are likely to find the show confusing, meaning small but significant proportion has expressed their difficulty on understanding the show.

\newpage

# PROBLEM 4: Advertisement on Ebay's Revenue Improvement

```{r, echo = FALSE, message = FALSE, warning = FALSE}
ebay <- read.csv("ebay.csv")

# ration of each DMA
ebay <- ebay |>
  mutate(revenue_ratio = rev_after / rev_before)

set.seed(111) # for reproducibility

# Bootstrap resampling for diff in mean revenue ratio btw treatment and control groups
boot_diff_rev <- do(10000) * diffmean(revenue_ratio ~ adwords_pause, data = resample(ebay))

# distribution has any varaibility?
ggplot(boot_diff_rev) +
  geom_histogram(aes(x=diffmean), binwidth = 0.01)+
  labs(
    title = "Bootstrap Distribution of Revenue Ratio Difference",
    x = "Difference in Mean Revenue Ratio",
    y = "Frequency"
  )

# calculate sd error
sd(~diffmean, data = boot_diff_rev)

# 95% confidence interval 
confint(boot_diff_rev, level = 0.95)

```

## Write-Up

-   Question: Is there a significant difference in the revenue rations between DMAs that paused paid advertising (treatment group) and DMAs that continued advertising (control group)? Is the treatment group has a lower revenue ratio, which would suggest the paid advertisements on Google contribute positively to EBay's revenue?
-   Approach: I used bootstrap resampling with n = 10000 to calculate difference in revenue ratio between treatment and control groups. Then I calculated a 95% confidence interval in percentile format.
-   Results: The 95% confidence interval for the difference in revenue ratios between treatment and control DMAs is (-0.0913, -0.0136), with the difference in the revenue ratio (original estimate point) of -0.0523 and SD error of 0.0199.
-   Conclusion: Since the entire interval is negative, this suggests that the treatment group (DMAs who paused paid ads) consistently had a lower revenue ratio. This infers that eBay's paid search advertising likely contributes positively to evenue as DMAs with active advertising had a higher revenue ratio than those who weren't.
